---
title: "External SDKs and database drivers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{External SDKs and database drivers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Obtain a connection from the module and ask it for a token immediately before
calling a library that manages its own HTTP or database transport:

```{r, eval = FALSE}
connection <- shiny::req(auth$connection())
token <- connection$access_token(required_scopes = "records.read")
existing_sdk_read_records(access_token = token)
```

The connection checks the owning Shiny session, authorization, scopes and
remaining lifetime. If needed, it performs one coordinated refresh and checks
the committed result again. It does not start login or retry the SDK operation.
Configure the client for the API's token audience and scopes; a valid token
string alone does not prove access to a particular resource.

## A single-authorization app

This complete app skeleton uses `httr2` as the external library. Replace the
example endpoints, scopes and registered callback URL with your provider's
values, and set `OAUTH_CLIENT_ID` and `OAUTH_CLIENT_SECRET` in the environment.
The token and authenticated connection stay inside `server()`.

```{r, eval = FALSE}
library(shiny)
library(shinyOAuth)

provider <- oauth_provider(
  name = "records",
  auth_url = "https://login.example/authorize",
  token_url = "https://login.example/token",
  userinfo_required = FALSE
)
client <- oauth_client(
  provider,
  client_id = Sys.getenv("OAUTH_CLIENT_ID"),
  client_secret = Sys.getenv("OAUTH_CLIENT_SECRET"),
  redirect_uri = "https://app.example/callback",
  scopes = c("records.read", "records.write"),
  required_scopes = "records.read"
)

ui <- oauth_ui(fluidPage(
  actionButton("login", "Sign in"),
  actionButton("again", "Sign in again"),
  actionButton("logout", "Sign out"),
  actionButton("load", "Load records"),
  textOutput("permissions"),
  verbatimTextOutput("records")
), id = "auth", client = client)

server <- function(input, output, session) {
  auth <- oauth_module_server(
    "auth", client, auto_redirect = FALSE, refresh_proactively = TRUE
  )
  observeEvent(input$login, {
    auth$request_login()
    invisible(NULL)
  })
  observeEvent(input$again, {
    auth$reauthorize()
    invisible(NULL)
  })
  observeEvent(input$logout, auth$logout())

  output$permissions <- renderText({
    connection <- req(auth$connection())
    if (connection$has_scopes("records.write")) "Editing available" else "Read only"
  })

  records <- eventReactive(input$load, {
    connection <- req(auth$connection())
    data <- tryCatch({
      response <- httr2::request("https://api.example/records") |>
        httr2::req_auth_bearer_token(
          connection$access_token(required_scopes = "records.read")
        ) |>
        httr2::req_perform()
      httr2::resp_body_json(response)
    }, shinyOAuth_access_error = function(error) {
      showNotification("Authorization unavailable. Try again or sign in again.")
      req(FALSE)
    })
    list(authorization = connection$id, data = data)
  })

  output$records <- renderPrint({
    connection <- req(auth$connection())
    result <- records()
    req(identical(result$authorization, connection$id))
    req(connection$has_scopes("records.read"))
    result$data
  })
}

shinyApp(ui, server, uiPattern = ".*")
```

`uiPattern = ".*"` lets Shiny route the registered `/callback` path through
`oauth_ui()` as well as serving the application root.

The output guard matters: `eventReactive()` retains its previous result until
the next click. Checking the current authorization ID and permissions prevents
that result from remaining visible after logout or a different login. Use the
same guard for cached or background results. Clear application caches when no
longer needed, and include the authorization ID in any session-local cache key.
Never share user data through a global cache keyed only by a query string.

`auth$connection()` returns `NULL` before login and after the authorization is
cleared. During one authorization it returns the same object and ID across
refresh. Old references remain invalid after replacement, even if the same
provider account signs in again. Token rotation alone does not rerun consumers
of the factory, `$access_token()` or `$has_scopes()`; explicit status reads such
as `$summary()` have finer reactive dependencies.

The single module's existing expiry and session-age policies still apply.
Proactive refresh helps keep a live authorization ready. With
`indefinite_session = TRUE`, the module retains expired tokens, so the accessor
can refresh them on demand; that option also disables automatic maximum-age
clearing. Neither option makes an expired token usable without refresh.

Automatic authentication-age expiry ends the local authorization. A subsequent
ordinary login requests the original configured scopes again. For example, after
`connection$refresh(scopes = "records.read")` narrows a `records.read records.write`
grant, ordinary login can request both permissions again. This also applies to
token targets. Explicit `auth$reauthorize()` preserves the retained narrower
limits. Configure only permissions the application should request at every fresh
login; local narrowing does not change the client configuration or revoke the
provider's underlying consent.

## SDK callbacks and databases

If an SDK accepts a token callback, create its client once per authorization:

```{r, eval = FALSE}
sdk <- reactive({
  connection <- req(auth$connection())
  example_sdk_client(
    token = function() connection$access_token(required_scopes = "records.read")
  )
})
```

Here `example_sdk_client()` represents your SDK's constructor. The callback
captures the authorization-bound connection, not a token snapshot or a lookup
that silently follows future logins. Call it in the owning session's reactive
context. It cannot be serialized to a worker or invoked from another session.

For a driver that accepts only a string, retrieve it immediately before opening
or updating the driver connection and follow that driver's cleanup rules. A
copied token does not refresh itself. Do not put tokens, SDK clients or database
connections in global app state, logs, URLs or browser outputs. Token-only
connections do not need `resource_bases`; shinyOAuth's `$request()` still rejects
every undeclared destination.

Run inserts and other writes from explicit input events. Check operation scopes
at submission, acquire a current token, and handle the API's own access checks.
`$has_scopes()` only checks the recorded grant; it does not prove a database role,
row permission or continued provider authorization. shinyOAuth never replays a
business request after refresh or reauthorization.

## Refresh, async work and recovery

`$access_token(min_valid_for = 60)` requires a known lifetime longer than the
requested buffer, including your allowance for clock skew. `force_refresh = TRUE`
bypasses the token cache but retains refresh coordination and retry pacing.
A replacement that is still too short-lived fails without another refresh loop.
`$refresh(scopes = ...)` can narrow the grant while keeping required permissions;
later refreshes retain that accepted limit. `required_scopes` on `$access_token()`
only checks permissions and does not request different scopes.

```{r, eval = FALSE}
# Configure mirai daemons or a non-sequential future plan before starting Shiny.
# Inside the owning Shiny session:
connection <- req(auth$connection())
pending <- connection$access_token(required_scopes = "records.read", async = TRUE)
promises::then(pending, function(token) {
  # Recheck before starting work after an asynchronous boundary.
  current <- req(auth$connection())
  req(identical(current$id, connection$id))
  req(current$has_scopes("records.read"))
  existing_sdk_read_records(access_token = token)
})
```

The async accessor always returns a promise, including cached results. This
only makes credential acquisition asynchronous; the SDK operation needs its
own async transport if it would block Shiny. Concurrent async acquisitions join
the same owned refresh. A synchronous call never waits on that work or returns
a promise. Validate ownership again before displaying later SDK results.
Scope narrowing keeps the connection ID, so also recheck the operation's
permissions before starting SDK work after an asynchronous boundary.

Authentication failures inherit `shinyOAuth_access_error` and expose
`error$context$reason`:

| Reason | Meaning |
| --- | --- |
| `authorization_unavailable` | The reference, authorization or owning session ended. |
| `insufficient_scope` | The recorded grant does not cover the operation. |
| `refresh_pending` | A synchronous caller cannot join an async refresh. |
| `refresh_unavailable` | Refresh failed or retry pacing prevents another attempt. |
| `interaction_required` | A new authorization is needed. |
| `lifetime_unavailable` | The refreshed token still lacks the requested lifetime. |
| `unsupported_token_binding` | A bearer string cannot supply the token's binding. |

Use the reason to present your own recovery message; argument and configuration
errors remain separate. DPoP and certificate-bound tokens need their configured
transport, such as `$request()`, instead of token export.

`auth$reauthorize()` invalidates the old local authorization immediately and
starts replacement with its retained scope limit. It does not revoke the
upstream grant, force a password/account chooser, or restore the old reference
if the user cancels. Existing `request_login()` and `logout()` retain their
behavior. No token accessor redirects the user automatically.

## Several authorizations and optional UserInfo

With `oauth_connections_server()`, obtain the same connection interface through
`auth$connection(connection_id)`. `auth$reauthorize(connection_id)` replaces that
authorization. Find the replacement in `auth$connections()` by its
`replaces_connection_id`, then select its new ID. Unrelated grants remain usable;
existing protections for shared refresh credentials still apply. Credential
acquisition does not extend owner inactivity limits. Call `auth$touch()` from
explicit user events where appropriate. See `vignette("multiple-authorizations")`.

If selection is unnecessary, `auth$connection()` returns the sole retained
authorization, or NULL when none remain. With several it raises
`shinyOAuth_access_error` with reason `selection_required`. Expired and uncertain
authorizations still count; the factory does not guess which account you intend.

For Microsoft API tokens that cannot call Graph UserInfo, use
`oauth_provider_microsoft(tenant = "organizations", userinfo_required = FALSE)`.
This skips profile retrieval while preserving the preset's OIDC validation.
Request the appropriate OIDC scopes and use `connection$identity()` for selected
validated ID-token claims. Generic `oauth_provider()` and OIDC configuration
already support optional UserInfo. For one authorization with several API tokens,
see `vignette("token-targets")`. With targets, `identity()` retains the last
validated identity independently of access-token expiry or current `openid`
scope; it does not prove fresh authentication or API access. Ordinary connections
require a usable access token with `openid` currently granted. Query and database
behavior belongs to the consuming library.

## Background tasks and result ownership

Read Shiny reactives and acquire the credential before invoking an
`ExtendedTask` (requires Shiny 1.8.1 or newer). Its function runs outside the reactive graph. Pass plain inputs
and a short-lived bearer token to the worker, never the connection reference,
module state or refresh credential. This follows Shiny's
[ExtendedTask contract](https://shiny.posit.co/r/articles/improve/nonblocking/).

```{r, eval=FALSE}
# Configure future::plan(future::multisession) before starting the app.
# Inside server(), with auth already created:
load_records <- shiny::ExtendedTask$new(function(token, authorization_id) {
  promises::future_promise({
    list(
      authorization_id = authorization_id,
      data = existing_sdk_read_records(access_token = token)
    )
  })
})

shiny::observeEvent(input$load, {
  shiny::req(load_records$status() != "running")
  connection <- shiny::req(auth$connection())
  token <- connection$access_token("records.read", min_valid_for = 120)
  load_records$invoke(token, connection$id)
})

output$records <- shiny::renderTable({
  connection <- shiny::req(auth$connection())
  shiny::req(connection$has_scopes("records.read"))
  result <- load_records$result()
  shiny::req(identical(result$authorization_id, connection$id))
  result$data
})
```

Choose a lifetime buffer that covers worker startup, any queue delay and the
operation. Do not queue these token snapshots behind long tasks; reacquire near
dispatch if your scheduler delays work. A copied bearer token cannot refresh
itself. A background failure does not automatically resubmit the operation.

The reactive ownership and permission checks clear displayed data on logout,
replacement, or removal of `records.read`, and reject a late result after any
of those changes. Scope narrowing keeps the authorization ID, so check both.
Keep these checks even when
the task was started by an input event. Do not use `cancelOutput = TRUE` here,
because it retains previously displayed data. For a manager with several
authorizations, apply the same check to the explicitly selected connection ID.
