---
title: "Keep and use multiple OAuth authorizations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Keep and use multiple OAuth authorizations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

shinyOAuth allows you to keep multiple OAuth authorizations in one Shiny app.
Users can connect several services, choose which saved authorization to use for
a request, and disconnect each one separately.

This is useful for a dashboard that combines data from several APIs, or an app
where users need access to several accounts at the same service.

This vignette builds an app that connects to two services and reads data from
either one. It uses browser retention to keep both authorizations available
when the user returns from a provider's login page.

## Client configuration and connections

Configure each service with an `OAuthClient` and give it a name in a `clients`
list. The connection manager stores each authorization separately. In the Shiny
server, select an `OAuthConnection` to make requests with that authorization.

| Component | Purpose |
| --- | --- |
| `OAuthClient` | Shared settings for a service: registration, scopes and approved API addresses. |
| `clients <- list(a = client_a, b = client_b)` | Local names used by the app's Connect buttons. |
| `oauth_connections()` | Shared configuration for storing and managing the authorizations. |
| `OAuthConnection` | A session-bound handle for making requests with one authorization's current credentials. |

One client can have several connections. Connecting A again creates another
authorization alongside the first. Account selection happens at the provider;
it may reuse the account already signed in there.

For apps that need one authorization per Shiny session, see the `oauth_ui()`
and `oauth_module_server()` example in [Usage](usage.html).

## Register two clients

This template uses two fictional services. Replace their endpoints, credentials,
scope names and API path with values from your providers' documentation. It
assumes authorization code flow with S256 PKCE and confidential clients using
HTTP Basic client authentication. The providers must support that registration.

Register these exact redirect URIs with the respective services:

| Service | Redirect URI |
| --- | --- |
| A | `http://127.0.0.1:8100/callback/a` |
| B | `http://127.0.0.1:8100/callback/b` |

These loopback HTTP addresses are for local development, if your providers
permit them. Use your public HTTPS origin and corresponding registered
callbacks when deploying, and omit the owner's `allow_http_loopback` exception.
Store the two client IDs and secrets in the environment
variables used below, outside source control.

Copy the next three code blocks, in order, into `app.R`.

### 1. Configure clients outside `server()`

```{r multiple-clients, eval = FALSE}
library(shiny)
library(shinyOAuth)

app_origin <- "http://127.0.0.1:8100"
callbacks <- paste0(app_origin, c("/callback/a", "/callback/b"))

provider_a <- oauth_provider(
  name = "Service A",
  auth_url = "https://login.a.example/authorize",
  token_url = "https://login.a.example/token",
  token_auth_style = "header",
  use_pkce = TRUE,
  pkce_method = "S256"
)
provider_b <- oauth_provider(
  name = "Service B",
  auth_url = "https://login.b.example/authorize",
  token_url = "https://login.b.example/token",
  token_auth_style = "header",
  use_pkce = TRUE,
  pkce_method = "S256"
)

client_a <- oauth_client(
  provider_a,
  client_id = Sys.getenv("SERVICE_A_CLIENT_ID"),
  client_secret = Sys.getenv("SERVICE_A_CLIENT_SECRET"),
  redirect_uri = callbacks[[1]],
  scopes = "records.read",
  resource_bases = c(api = "https://api.a.example/v1"),
  required_scopes = "records.read",
  label = "Service A",
  authorization_server_mode = "multi_redirect_uri",
  authorization_server_redirect_uris = callbacks
)
client_b <- oauth_client(
  provider_b,
  client_id = Sys.getenv("SERVICE_B_CLIENT_ID"),
  client_secret = Sys.getenv("SERVICE_B_CLIENT_SECRET"),
  redirect_uri = callbacks[[2]],
  scopes = "records.read",
  resource_bases = c(api = "https://api.b.example/v1"),
  required_scopes = "records.read",
  label = "Service B",
  authorization_server_mode = "multi_redirect_uri",
  authorization_server_redirect_uris = callbacks
)

clients <- list(a = client_a, b = client_b)
```

`a` and `b` are app-local selectors, independent of the registered `client_id`.
`api` names an approved resource within each client. For connection A,
`$request("api", "records")` means `https://api.a.example/v1/records`.

The default callback policy gives each client a distinct route. Supply the
**complete callback vector to each client**, even though each uses its own
`redirect_uri`. This lets the package check the routing configuration. Shared
callback routes are an advanced alternative documented in `?oauth_connections`.

### 2. Keep authorizations when the browser returns

```{r multiple-manager, eval = FALSE}
# Temporary keys for this local demo, generated once per R process.
# In a deployment, load two independent 32-byte raw keys from secret storage.
keys <- list(
  credentials = openssl::rand_bytes(32),
  owner = openssl::rand_bytes(32)
)

manager <- oauth_connections(
  clients,
  app_origin = app_origin,
  retention = "browser",
  owner = oauth_browser_owner(allow_http_loopback = TRUE),
  store = oauth_connection_store_memory(),
  keys = keys
)
```

Browser retention uses an HttpOnly cookie containing an opaque local owner ID.
Tokens stay encrypted in the R process. The cookie associates saved connections
with this browser; it does not identify a person or log them into either service.

Without `retention = "browser"`, the default is `"shiny"`: existing connections
are discarded when that Shiny session ends. Merely configuring multiple clients
does not preserve A while the browser goes away to authorize B.

Keep the manager, store and keys outside `server()` so returning Shiny sessions
use the same configuration. This memory store supports **one R process** and
does not survive an R restart. Stable keys alone do not make it persistent or
share records between workers. Browser ownership also has idle and absolute
expiry limits; refresh does not extend them. For an app with its own verified
account login, see `?oauth_account_owner` and `retention = "account"`.

### 3. Select a connection and make a request

```{r multiple-app, eval = FALSE}
base_ui <- fluidPage(
  use_shinyOAuth(),
  h2("My connected services"),
  actionButton("connect_a", "Connect Service A"),
  actionButton("connect_b", "Connect Service B"),
  selectInput("connection_id", "Saved authorization", choices = character()),
  actionButton("read", "Read records"),
  actionButton("disconnect", "Disconnect selected"),
  actionButton("logout", "Disconnect all and leave"),
  textOutput("result"),
  textOutput("auth_error")
)
ui <- oauth_connections_ui(base_ui, "services", manager)

server <- function(input, output, session) {
  auth <- oauth_connections_server("services", manager)
  observeEvent(input$connect_a, auth$connect("a"))
  observeEvent(input$connect_b, auth$connect("b"))

  observe({
    rows <- auth$connections()
    ids <- vapply(rows, function(x) x$connection_id, character(1))
    labels <- vapply(seq_along(rows), function(i) {
      paste(rows[[i]]$client_label, i, paste0("(", rows[[i]]$status, ")"))
    }, character(1))
    selected <- isolate(input$connection_id)
    if (!length(selected) || !selected %in% ids) selected <- head(ids, 1)
    updateSelectInput(session, "connection_id",
      choices = setNames(ids, labels), selected = selected)
  })

  selected_connection <- reactive({
    rows <- auth$connections()
    ids <- vapply(rows, function(x) x$connection_id, character(1))
    req(input$connection_id, input$connection_id %in% ids)
    auth$connection(input$connection_id)
  })

  result <- eventReactive(input$read, {
    connection <- selected_connection()
    req(connection$is_usable())
    message <- tryCatch({
      response <- connection$request(
        "api", "records", required_scopes = "records.read"
      )
      httr2::resp_check_status(response)
      # Parse the body here according to your API's schema.
      paste("Records request succeeded; HTTP", httr2::resp_status(response))
    }, error = function(e) "Could not read records. Check the connection and try again.")
    list(connection_id = connection$id, message = message)
  })

  output$result <- renderText({
    connection <- selected_connection()
    req(connection$is_usable())
    value <- result()
    req(identical(value$connection_id, connection$id))
    value$message
  })
  observeEvent(input$disconnect, {
    auth$disconnect(selected_connection()$id)
  })
  observeEvent(input$logout, auth$logout())
  output$auth_error <- renderText({
    if (length(auth$errors())) "Authorization is unavailable. Try connecting again."
  })
}

runApp(shinyApp(ui, server, uiPattern = ".*"),
  host = "127.0.0.1", port = 8100, launch.browser = FALSE)
```

Open `http://127.0.0.1:8100` in a regular browser. Connect A, return, then connect
B. Both authorizations should appear in the selector. Select each and read its
records; disconnecting one leaves the other available. The numeric suffix only
distinguishes rows in this demo; use the opaque `connection_id` as the stable key.

The UI wrapper handles callback pages; `uiPattern = ".*"` lets Shiny serve those
paths. Use the same module ID (`"services"`) in UI and server. Deploying under a
subpath or behind a proxy can also require `app_base_path` and a trusted
`request_uri_resolver`; see `?oauth_connections_ui`.

## Requests, permissions and refresh

`OAuthConnection$request()` uses the existing `perform_resource_req()` transport.
It adds selection of the current owned credentials and checks the approved API
base and required permissions. Existing code using `perform_resource_req(token,
url)` remains supported. For a single existing module, `oauth_connection(client,
reactive(auth$token))` supplies the optional wrapper without adding retention.

`resource_bases` limits destinations, including absolute pagination links and
their base paths. It does not automatically send OAuth's `resource` parameter
or prove the audience of an opaque token; configure resource indicators
explicitly when your provider requires them. See [RFC 8707](https://www.rfc-editor.org/rfc/rfc8707.html).

`required_scopes` on the client is the minimum for usable access. Requesting
additional optional scopes can produce a `limited` connection if only some are
granted. Supply an operation's additional permissions in `$request()`; ordinary
OAuth scope names are compared literally, and permissions are not inferred from
the API path. A usable connection does not guarantee the API will accept every
request.

If the provider issues a refresh token, the manager attempts refresh at expiry;
`refresh_proactively = TRUE` on `oauth_connections_server()` starts earlier.
Explicit refresh is also available inside a server observer:

```{r multiple-refresh, eval = FALSE}
observeEvent(input$refresh, {
  tryCatch(selected_connection()$refresh(), error = function(e) {
    showNotification("Could not refresh. Connect again if access has ended.", type = "error")
  })
})
```

Add `actionButton("refresh", "Refresh selected")` to the UI to use this snippet.
The example uses the default synchronous mode. `async = TRUE` needs worker setup
and promise handling for explicit refresh; see [Usage](usage.html#asynchronous-execution).

Retention itself does not request a refresh token. OAuth refresh cannot expand
the authorization's scope ([RFC 6749, section 6](https://www.rfc-editor.org/rfc/rfc6749.html#section-6)).
`$refresh(scopes = ...)` can deliberately narrow a managed connection's accepted
permissions; getting broader access again requires a new authorization.

Disconnect immediately removes local usability. Remote token revocation is
best effort and reported separately in the return value. This template does not
configure revocation endpoints, so add each provider's documented
`revocation_url` if supported. `logout()` also ends this local owner session;
it does not sign the person out of the external services or your own account
login system.
