---
title: "Research and institutional provider setup"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Research and institutional provider setup}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

These presets configure providers; you still register an application with the
provider and supply its client ID, secret, and redirect URI. The recipes below
use confidential web applications with a server-side secret. Except for
Globus, the helpers fetch discovery metadata: construct the provider once
during app startup, outside `server()`.

## Choose a provider

| Service | Provider construction | Initial login scopes |
|---|---|---|
| ORCID | `oauth_provider_orcid()` | `"openid"` |
| GitLab | `oauth_provider_gitlab()` | `c("openid", "profile", "email")` |
| Amazon Cognito | `oauth_provider_cognito(issuer)` | `c("openid", "profile", "email")` |
| Globus Auth | `oauth_provider_globus()` | `c("openid", "profile", "email")` |
| Hugging Face | `oauth_provider_huggingface()` | `c("openid", "profile")` |
| SURFconext | `oauth_provider_surfconext()` | `"openid"` |
| authentik | `oauth_provider_authentik(base_url, application_slug)` | `c("openid", "profile", "email")` |

All seven presets require and validate an ID token and verify that the UserInfo
subject matches it. Request only the profile claims you need. Scopes belong on
`oauth_client()`, not on the provider helper. All presets use S256 PKCE except
the ORCID confidential-client preset, which retains client authentication and
OIDC nonce validation but does not claim PKCE support.

## A Shiny app for any recipe

Register the app's exact redirect URI with your provider. Use a deployed HTTPS
address for production. For providers that permit local development, the
example below uses `http://127.0.0.1:8100`; ORCID production requires HTTPS.
Set `OAUTH_CLIENT_ID`, `OAUTH_CLIENT_SECRET`, and `OAUTH_REDIRECT_URI` in your
environment, then save this as `app.R`. Replace the provider and scopes with
one recipe below.

```{r, eval=FALSE}
library(shiny)
library(shinyOAuth)

provider <- oauth_provider_gitlab()
client <- oauth_client(
  provider = provider,
  client_id = Sys.getenv("OAUTH_CLIENT_ID"),
  client_secret = Sys.getenv("OAUTH_CLIENT_SECRET"),
  redirect_uri = Sys.getenv("OAUTH_REDIRECT_URI"),
  scopes = c("openid", "profile", "email")
)

ui <- oauth_ui(
  fluidPage(h2("Provider login"), textOutput("identity")),
  id = "auth",
  client = client
)

server <- function(input, output, session) {
  auth <- oauth_module_server("auth", client)
  output$identity <- renderText({
    req(auth$authenticated)
    paste("Signed in as", auth$token@userinfo[["sub"]])
  })
}

shinyApp(ui, server, options = list(port = 8100))
```

Open the app in a regular browser. The module handles state, the redirect,
code exchange, and identity validation. Keep credentials and complete token
objects out of the UI and logs. For application account keys, use the validated
issuer and subject together. Successful login alone does not decide which
application resources that account may access.

## ORCID

1. Register a Public API client through **Developer Tools** in your ORCID
   account, or arrange a Member API client with your organization.
2. Register the redirect URI. Sandbox users and credentials are separate from
   production; use `oauth_provider_orcid(sandbox = TRUE)` for testing.
3. Request `scopes = "openid"`. ORCID says to replace `/authenticate` with
   `openid` for OIDC, rather than request both. Do not copy `profile` and
   `email` scopes from another provider's example.

The helper selects body-based client authentication. ORCID's discovery does
not advertise PKCE, so the preset uses the confidential authorization-code
flow with an OIDC nonce. ORCID record read/update permissions are separate
from identifying a researcher and may require Member API credentials.

Source: [ORCID authentication guide](https://info.orcid.org/documentation/api-tutorials/api-tutorial-get-and-authenticated-orcid-id/).

## GitLab

1. Create an OAuth application in your GitLab instance's application settings.
   Register the callback and enable the OIDC scopes needed by your app.
2. Use `oauth_provider_gitlab()` for GitLab.com, or
   `oauth_provider_gitlab("https://gitlab.example.edu")` for your instance.
3. Request `openid`, with `profile` and `email` if needed. Request GitLab API
   or repository scopes separately when the app actually uses those APIs.

Discovery infers the token authentication method. If your registration needs
a particular advertised method, supply `token_auth_style`, for example
`"body"` for `client_secret_post`.

Source: [GitLab as an OIDC provider](https://docs.gitlab.com/integration/openid_connect_provider/).

## Amazon Cognito

1. Create or select a user pool and configure its managed-login/custom domain.
2. Create an app client with a secret, enable the authorization code grant and
   required OIDC scopes, and register the callback URI.
3. Copy the pool's exact issuer URL into the helper:

```{r, eval=FALSE}
provider <- oauth_provider_cognito(
  "https://cognito-idp.eu-west-1.amazonaws.com/eu-west-1_Example"
)
```

The issuer is not the login domain. Discovery returns the login-domain
endpoints while tokens are validated against the user pool issuer. If AWS
configures the pool with the updated `issuer-cognito-idp` issuer form, supply
that exact URL instead. Supplying the URL also avoids assuming an AWS partition
or hostname suffix. Enable `openid`, `profile`, and `email` as needed in both
the app registration and `oauth_client()`.

Source: [Cognito federation endpoints](https://docs.aws.amazon.com/cognito/latest/developerguide/federation-endpoints.html).

## Globus Auth

1. Register a confidential web application in the
   [Globus developer settings](https://app.globus.org/settings/developers).
2. Configure its redirect URI and client secret.
3. Use `oauth_provider_globus()` and login scopes
   `c("openid", "profile", "email")`. Add `offline_access` only when a
   refresh token is needed.

Globus advertises RS512-only ID token signing. The preset therefore uses its
published endpoints and JWKS URL directly, with RS512 validation, instead of
loosening generic discovery's RS256 requirement.

**This preset covers OIDC login.** Globus can issue several access and refresh
tokens for different APIs in a single response. Additional credentials are
preserved in `token@extra_fields[["other_tokens"]]`, but are not automatically
selected, refreshed, or revoked by shinyOAuth. Restrict this login client to
the OIDC scopes above. For Transfer, Compute, or other multi-resource grants,
use a separate integration that manages each resource server's credentials.
Access granted to one resource server does not authorize sending its token
to another server. Extra fields are credentials, not validated ID token claims.

Source: [Globus Auth developer guide](https://docs.globus.org/api/auth/developer-guide/).

## Hugging Face

1. Register an OAuth app in your Hugging Face account settings, retaining a
   client secret for this confidential Shiny app.
2. Register the redirect URI and use `oauth_provider_huggingface()`.
3. Start with `c("openid", "profile")`; request `email` only if needed.

Hub and inference permissions are separate: `read-repos` authorizes repository
reads, while `inference-api` authorizes inference calls on the user's behalf.
Register and request the additional permissions your app needs, and configure
approved API resource bases on the client before making resource requests.

Source: [Sign in with Hugging Face](https://huggingface.co/docs/hub/oauth).

## SURFconext

1. Arrange service onboarding with SURFconext and register the callback URI.
   Agree which institutional attributes your service needs.
2. Use `oauth_provider_surfconext(environment = "test")` for the test service,
   then `oauth_provider_surfconext()` with production credentials.
3. Request `"openid"`. Additional scope requests do not override the agreed
   attribute release policy. Do not assume that email or a name is present.

Use the validated issuer and subject as the account identifier. Some services
receive only pseudonymous identifiers and affiliation information.

Source: [SURFconext service-provider documentation](https://servicedesk.surf.nl/wiki/spaces/IAM/pages/128909810/SURFconext+for+Service+Providers).

## authentik

1. Create an OAuth2/OIDC provider and attach it to an application. Configure
   the callback, a confidential client, and an RS256-capable signing key.
2. Keep the default **per-application issuer** mode, and copy the application
   slug (not its display name):

```{r, eval=FALSE}
provider <- oauth_provider_authentik(
  "https://auth.example.com", application_slug = "shiny-app"
)
```

3. Enable the scope mappings for `openid`, `profile`, and `email` as needed.
   For refresh tokens, enable the `offline_access` mapping and request that
   scope on the client.

The helper preserves the trailing slash in the application issuer. Global
issuer mode uses a different issuer identifier and must be configured manually
with `oauth_provider()`; this preset rejects that mismatch.

Source: [authentik OAuth2/OIDC provider](https://docs.goauthentik.io/add-secure-apps/providers/oauth2/).
