---
title: "Use SMART on FHIR from Shiny"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Use SMART on FHIR from Shiny}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

shinyOAuth allows your Shiny app to request access to a FHIR server using SMART
on FHIR, read data for the selected patient, and keep separate authorizations
for different healthcare providers.

This is useful for a patient dashboard that reads records from a hospital, or
a clinical app that opens from an electronic health record (EHR) with a patient
already selected.

This vignette builds an app that connects to a hospital and reads a Patient
resource. It also shows how to search for Observations, identify the logged-in
user, and configure the app to launch from an EHR.

## FHIR and SMART on FHIR

**FHIR** defines healthcare data structures and ways to exchange them. A
`Patient` resource describes a patient; an `Observation` can describe a clinical
measurement. An address such as `https://ehr.example/fhir/R4` is the base of a
FHIR API. See the [FHIR overview](https://hl7.org/fhir/R4/overview.html).

**OAuth** lets an app obtain permission to call an API. **SMART on FHIR** specifies
how to use OAuth with FHIR: discover authorization endpoints, request healthcare
permissions, identify the intended FHIR server, and obtain launch context such
as the selected patient. The EHR can open the app, or the person can open it
independently. See [SMART App Launch](https://hl7.org/fhir/smart-app-launch/STU2.2/app-launch.html).

The example below uses `smart_discover()` to read the server's SMART metadata
and `smart_client()` to configure your registered app. After the user authorizes
access, `smart_patient()` reads the selected patient's record using the saved
connection.

To connect several healthcare providers, pass a named list of clients to the
connection manager, as shown in
[Keep and use multiple OAuth authorizations](multiple-authorizations.html).

## Before running the example

Register the app with the chosen server. Obtain its exact FHIR base, client ID,
authentication method, and permitted scopes. This example uses a **standalone
confidential client with a shared secret** (`token_auth_style = "header"`). A
server-side Shiny deployment can keep that secret outside the browser.

The example's `.example` endpoints are placeholders, not a running sandbox.
Replace them with your approved server's addresses. Register
`http://127.0.0.1:8100/callback/fhir` for local development if your provider permits
loopback HTTP, and set `SMART_CLIENT_ID` and `SMART_CLIENT_SECRET` in your R
environment. Use HTTPS and matching registered URLs for deployment, and omit
both `allow_http_loopback` exceptions shown below.

This package validates a SMART App Launch STU 2.2 profile, including advertised
launch, authentication and scope capabilities. Discovery does not register your
app or guarantee that the server will grant its requested access.

Copy the next three code blocks, in order, into `app.R`.

### 1. Discover the server and configure the client

```{r smart-configuration, eval = FALSE}
library(shiny)
library(shinyOAuth)

app_origin <- "http://127.0.0.1:8100"
discovery <- smart_discover(
  "https://ehr.example/fhir/R4",
  endpoint_hosts = c("ehr.example", "login.example"),
  allow_http_loopback = TRUE # Enable the registered local HTTP callback for this demo.
)

client <- smart_client(
  discovery,
  client_id = Sys.getenv("SMART_CLIENT_ID"),
  client_secret = Sys.getenv("SMART_CLIENT_SECRET"),
  token_auth_style = "header",
  redirect_uri = paste0(app_origin, "/callback/fhir"),
  launch = "standalone",
  scopes = c("launch/patient", "patient/Patient.r", "patient/Observation.s"),
  required_scopes = "patient/Patient.r",
  label = "My hospital"
)
```

Discovery runs outside `server()`, against a deployment-configured base.
`endpoint_hosts` must include every approved hostname used by the metadata,
including a separate authorization server when applicable. Its default permits
only the FHIR base's hostname. Review these addresses and metadata when setting
up or updating the app; do not discover arbitrary URLs supplied by visitors.

`smart_client()` configures S256 PKCE, sets `aud` to this exact FHIR base, and
adds `resource_bases = c(fhir = ...)` to the client. You do not manually build the
authorization URL or copy the access token into a request.

The requested scopes have separate jobs:

| Scope | Why request it? |
| --- | --- |
| `launch/patient` | Ask the standalone authorization flow to establish patient context. |
| `patient/Patient.r` | Read the contextual patient's Patient resource. |
| `patient/Observation.s` | Search Observations in the authorized patient context. Optional in this example. |

In SMART v2 syntax, `.r` means read, `.s` means search, and `.rs` means both.
The patient context and the permission to read data are distinct. The server
still decides what access to allow. See [SMART scopes and launch context](https://hl7.org/fhir/smart-app-launch/STU2.2/scopes-and-launch-context.html).

Here `required_scopes` allows the app to work if Observation search is declined.
Such a connection is `limited`; Patient reading can still succeed. Omitting
`required_scopes` would make all requested scopes required. Identity and offline
access are not requested by this example.

### 2. Configure retained connections

```{r smart-manager, eval = FALSE}
manager <- oauth_connections(
  clients = list(hospital = client),
  app_origin = app_origin,
  retention = "browser",
  owner = oauth_browser_owner(allow_http_loopback = TRUE),
  store = oauth_connection_store_memory(),
  # Temporary keys for this local demo, created once at app startup.
  # For deployment, load independent 32-byte raw keys from secret storage.
  keys = list(
    credentials = openssl::rand_bytes(32),
    owner = openssl::rand_bytes(32)
  )
)
```

`hospital` is your local configuration name; it is unrelated to the registered
client ID or the patient's ID. Browser retention lets the app restore saved
grants after authorization redirects. It uses an HttpOnly owner cookie and
encrypted storage in **one R process**. Records disappear when that process
restarts; independent workers do not share this store. The
[multiple-authorizations vignette](multiple-authorizations.html#keep-authorizations-when-the-browser-returns)
explains ownership, expiry and deployment constraints.

### 3. Authorize and fetch the patient

```{r smart-app, eval = FALSE}
base_ui <- fluidPage(
  use_shinyOAuth(),
  h2("FHIR patient example"),
  actionButton("connect", "Connect hospital"),
  selectInput("connection_id", "Saved authorization", choices = character()),
  actionButton("load_patient", "Load patient"),
  actionButton("disconnect", "Disconnect selected"),
  textOutput("patient"),
  textOutput("auth_error")
)
ui <- oauth_connections_ui(base_ui, "health", manager)

server <- function(input, output, session) {
  health <- oauth_connections_server("health", manager)
  observeEvent(input$connect, health$connect("hospital"))

  observe({
    rows <- health$connections()
    ids <- vapply(rows, function(x) x$connection_id, character(1))
    labels <- vapply(seq_along(rows), function(i) {
      paste(rows[[i]]$client_label, i, paste0("(", rows[[i]]$status, ")"))
    }, character(1))
    selected <- isolate(input$connection_id)
    if (!length(selected) || !selected %in% ids) selected <- head(ids, 1)
    updateSelectInput(session, "connection_id",
      choices = setNames(ids, labels), selected = selected)
  })

  selected_connection <- reactive({
    rows <- health$connections()
    ids <- vapply(rows, function(x) x$connection_id, character(1))
    req(input$connection_id, input$connection_id %in% ids)
    connection <- health$connection(input$connection_id)
    req(connection$is_usable())
    connection
  })

  patient <- eventReactive(input$load_patient, {
    health$touch()
    connection <- selected_connection()
    context <- smart_context(connection)
    data <- tryCatch({
      response <- smart_patient(connection)
      httr2::resp_check_status(response)
      httr2::resp_body_json(response, simplifyVector = FALSE)
    }, error = function(e) NULL)
    list(connection_id = connection$id, revision = context$revision, data = data)
  })

  output$patient <- renderText({
    connection <- selected_connection()
    context <- smart_context(connection)
    value <- patient()
    # Clear the display if the selected grant or accepted context has changed.
    req(identical(value$connection_id, connection$id),
        identical(value$revision, context$revision))
    validate(need(!is.null(value$data), "Could not load this patient's record."))
    paste("Patient resource:", value$data$id)
  })
  observeEvent(input$disconnect, {
    # Allow disconnect even when the connection has become unusable.
    req(input$connection_id)
    health$disconnect(input$connection_id)
  })
  output$auth_error <- renderText({
    if (length(health$errors())) "Authorization is unavailable. Try authorizing again."
  })
}

runApp(shinyApp(ui, server, uiPattern = ".*"),
  host = "127.0.0.1", port = 8100, launch.browser = FALSE)
```

Open `http://127.0.0.1:8100` in a regular browser, connect the hospital, complete
its authorization flow, then load the patient. The helper reads the Patient
identified by the accepted context using the same connection's credentials and
configured FHIR base. It returns an `httr2` response.

This teaching UI displays only the resource ID. A clinical application needs
appropriate patient identification and must respect `need_patient_banner` when
provided. Context and resource bodies contain sensitive data; keep them out of
general connection tables and logs. Key patient-dependent caches by both
connection ID and `context$revision`, and discard stale data when either changes.

## Search with the same connection

Inside a server observer, use the existing request interface for other FHIR
operations. Add `actionButton("search", "Search observations")` to the UI to
use this example:

```{r smart-search, eval = FALSE}
observeEvent(input$search, {
  health$touch()
  connection <- selected_connection()
  context <- smart_context(connection)
  req(context$patient)
  tryCatch({
    response <- connection$request(
      "fhir", "Observation",
      query = list(patient = context$patient, `_count` = 10),
      required_scopes = "patient/Observation.s"
    )
    httr2::resp_check_status(response)
    bundle <- httr2::resp_body_json(response, simplifyVector = FALSE)
    # Process bundle$entry here, associating data with this ID and revision.
    showNotification("Observation search completed.")
  }, error = function(e) {
    showNotification("Could not search observations. Check the granted permissions.",
      type = "error")
  })
})
```

The `patient` query selects data; it does not enforce authorization. The server
enforces access, and the connection checks the declared operation scope. SMART
scope comparisons understand supported combinations such as `.rs` covering
`.r` and `.s`. The package does not infer permissions from arbitrary HTTP paths.
Pagination links must remain inside this connection's approved FHIR base.

Granular scopes support simple search parameter names: an optional leading
underscore, then a letter, followed by letters, digits or hyphens, with nonempty
values. This includes `_id`, `_lastUpdated`, `_tag`, `_profile` and `_security`.
Constraints are compared exactly as written; the package does not decode,
reorder or infer implication between searches. Modifiers, chaining and
experimental `_filter` expressions remain unsupported.

SMART refresh requests send the latest accepted scopes, including after the
server reduces a grant without rotating the refresh token. Refreshed grants may
retain or narrow those permissions. An explicitly empty grant remains valid,
but requires a fresh authorization to obtain more permissions; an empty OAuth
refresh scope cannot express that limit.

## Patient context and user identity are different

`smart_context(connection)$patient` identifies the chart in context. A clinician
can be the logged-in user while that patient is someone else. The browser owner
used for local retention is a third, separate concept.

If the app needs the authenticated user's FHIR reference, configure
`identity = "fhirUser"` when constructing its SMART client. This adds and
requires `openid fhirUser`, signed ID-token validation and nonce binding.
The server must advertise the corresponding capabilities and signing metadata.
`smart_context(connection)$fhirUser` then exposes the validated reference.

Fetching that resource with `smart_fhir_user(connection)` accepts the granted
`openid fhirUser` identity scopes, as described by
[SMART identity semantics](https://hl7.org/fhir/smart-app-launch/STU2.2/scopes-and-launch-context.html#scopes-for-requesting-identity-data).
It also accepts a matching user-level read permission, or `patient/Patient.r`
when the identity refers to the contextual Patient. The server enforces its own
access rules. Absolute identity URLs need not use a REST-style resource path.
A user reference outside the configured FHIR base is not fetched with this token.

## Start from an EHR instead

For EHR launch, the EHR opens a registered entry URL carrying `iss` (the FHIR
base) and an opaque `launch` handle. The handle is exchanged through the
authorization flow; the initial URL does not establish trusted patient context.
See [EHR launch](https://hl7.org/fhir/smart-app-launch/STU2.2/app-launch.html#launch-app-ehr-launch).

To adapt the standalone app, replace its client construction with:

```{r smart-ehr-client, eval = FALSE}
client <- smart_client(
  discovery,
  client_id = Sys.getenv("SMART_EHR_CLIENT_ID"),
  client_secret = Sys.getenv("SMART_EHR_CLIENT_SECRET"),
  token_auth_style = "header",
  redirect_uri = paste0(app_origin, "/callback/fhir"),
  launch = "ehr",
  scopes = c("patient/Patient.r", "patient/Observation.s"),
  required_scopes = "patient/Patient.r",
  label = "My hospital"
)
```

Use the credentials for your EHR-launch registration. The constructor adds
`launch`; do not include the standalone `launch/patient` request. Recreate the
manager with this client, then replace the UI wrapper with:

```{r smart-ehr-ui, eval = FALSE}
ui <- oauth_connections_ui(base_ui, "health", manager,
  launch_routes = list(smart_launch_route("/smart/launch", "hospital")))
```

Register `http://127.0.0.1:8100/smart/launch` as the development launch URL, in
addition to the separate callback URL. For deployment register the corresponding
HTTPS URLs. Remove the Connect button and its observer: a fresh EHR launch now
starts authorization automatically. Keep the remaining server code and the
`uiPattern = ".*"` app construction.

The route selects only the configured `hospital` client by its exact FHIR base;
it does not discover a new server from the incoming `iss`. Each attempt uses a
fresh launch transaction. Calling `health$connect("hospital")` for an EHR client
reports `fresh_ehr_launch_required`; reconnect by opening the app from the EHR.
Keep launch handles out of shared client configuration and access-log queries.

Each browser owner can hold up to eight unconsumed launch tickets across all
manager routes; the manager retains at most 1,000 in total. Extra entries are
rejected without removing existing tickets. Consumption or expiry frees space.
Protect both launch entry and owner-creating application pages with ingress
rate limits at a trusted reverse proxy. Browser cookies do not identify a person
or network client, and callers can obtain additional owners, so these quotas
do not replace those deployment controls. Use the proxy's verified client
address for rate limits and size the owner registry for the admitted traffic.

The current EHR entry implementation requires **top-level browser navigation,
browser retention with a Lax owner cookie, and one R process**. Embedded iframe
launch is not supported. A route can allow multiple named clients with distinct
FHIR bases. Separate registrations for the same base need separate entry routes.

## More than one hospital

Create one reviewed discovery snapshot and `smart_client()` per registration,
then pass `list(hospital_a = client_a, hospital_b = client_b)` to the manager.
Use distinct registered callbacks and configure every client with
`authorization_server_mode = "multi_redirect_uri"` plus the full callback vector,
as in the [general vignette](multiple-authorizations.html#configure-clients-outside-server).
The selector and request code stay the same. Each connection carries its own
credentials, FHIR base and context; patient IDs from different bases are not
interchangeable. To support both standalone and EHR entry, configure separate
clients for those flows.

## Other registration and refresh options

For SMART asymmetric registrations, code exchange and refresh identify the
client through the signed assertion and omit form-body `client_id`. This follows
the [App Launch token parameter table](https://hl7.org/fhir/smart-app-launch/STU2.2/app-launch.html#obtain-access-token)
and its [asymmetric worked example](https://hl7.org/fhir/smart-app-launch/STU2.2/example-app-launch-asymmetric-auth.html#retrieve-access-token).
The signature-verification text also describes checking a supplied client ID;
this package chooses the table's request shape. Ordinary OAuth JWT client
authentication retains its existing serialization.

Choose authentication to match the server registration:

| Registration | `smart_client()` settings |
| --- | --- |
| Public | `token_auth_style = "public"`; omit the secret and assertion key. |
| Confidential, shared secret | `token_auth_style = "header"` and `client_secret`. |
| Confidential, asymmetric | `token_auth_style = "private_key_jwt"`, private key, key ID, and supported `RS384` or `ES384`. |

For example, replace the original client construction with this for an
asymmetric standalone registration. Register the matching public key and key ID
with the server first; keep the private key on the application server.

```{r smart-asymmetric, eval = FALSE}
client <- smart_client(
  discovery,
  client_id = Sys.getenv("SMART_ASYMMETRIC_CLIENT_ID"),
  redirect_uri = paste0(app_origin, "/callback/fhir"),
  launch = "standalone",
  scopes = c("launch/patient", "patient/Patient.r"),
  required_scopes = "patient/Patient.r",
  token_auth_style = "private_key_jwt",
  client_assertion_private_key = openssl::read_key(Sys.getenv("SMART_PRIVATE_KEY_FILE")),
  client_assertion_private_key_kid = Sys.getenv("SMART_KEY_ID"),
  client_assertion_alg = "RS384",
  label = "My hospital"
)
```

The server must advertise the chosen method and algorithm. The package creates
and signs the assertion; app code does not assemble JWTs. See
[SMART asymmetric client authentication](https://hl7.org/fhir/smart-app-launch/STU2.2/client-confidential-asymmetric.html).

If your use case needs access when the user is offline, request `offline_access`
only with an appropriate registration and advertised support. Saving a browser
connection by itself does not request offline access or a refresh token. The
manager refreshes when a refresh credential is available. SMART responses need
an explicit positive `expires_in`; the package does not guess their lifetime.

`connection$refresh(scopes = ...)` can narrow the connection's accepted scopes.
That local limit is retained and sent on later refreshes; it does not prove the
server has reduced the original refresh-token grant. Broader access requires a
new authorization. On refresh, read `smart_context()` again and use its revision
to invalidate stale patient data. Context tracking here does not subscribe to
live changes in the EHR's open chart.

Outgoing `authorization_method = "POST"` is available when discovery advertises
`authorize-post`. It is independent of `response_mode = "form_post"`, which
controls the returning callback. The default example uses GET authorization
and query callbacks. SMART composition with JAR, PAR, JARM, DPoP and mTLS is not
currently exposed through `smart_client()`; consult its reference for scope.

## Testing with fixtures and the Docker sandbox

The repository has browser integration tests for standalone and EHR launch,
multiple retained connections, supported registration types, refresh, context
and outgoing POST. Independent cryptographic conformance tests cover client
assertions. See the [coverage map](https://github.com/lukakoning/shinyOAuth/blob/2200f1bf7d44670a2c7f70a297b4cde787a3cc93/integration/smart/coverage.md)
for recorded results and runner commands.

The official `smart-on-fhir/smart-dev-sandbox` Docker setup is also included.
The pinned launcher advertises asymmetric authentication but omits required
signing-algorithm metadata, so it currently fails strict SMART discovery.
Passing its diagnostic checks does **not** establish a working external SMART
authorization flow. Use the [sandbox guide](https://github.com/lukakoning/shinyOAuth/blob/2200f1bf7d44670a2c7f70a297b4cde787a3cc93/integration/smart/sandbox.md)
for setup and the compatibility gate; external app interoperability remains an
open release checkpoint. Do not bypass metadata validation to make this example
appear compatible.
