---
title: "Several API tokens from one authorization"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Several API tokens from one authorization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(shinyOAuth)
```

A token target names the resource and API scopes for one access token. Several
targets can share a user's local authorization and refresh credential while
keeping separate access-token responses. This differs from several independent
authorizations, which continue to use `oauth_connections_server()` and separate
connection IDs.

The provider must explicitly support the acquisition protocol. A declaration is
a request limit, not evidence of consent. The server can refuse a target or omit
the refresh credential needed to acquire it. No target method starts sign-in,
switches accounts or tenants, or implements token exchange/OBO.

## Declare calendar and contacts targets

```{r, eval=FALSE}
# Use this mode only after confirming RFC 8707 support with your provider.
provider <- oauth_provider(
  "workplace", "https://login.example/authorize", "https://login.example/token",
  token_target_mode = "rfc8707",
  token_auth_style = "body",
  userinfo_required = FALSE
)
client <- oauth_client(
  provider,
  client_id = Sys.getenv("OAUTH_CLIENT_ID"),
  client_secret = Sys.getenv("OAUTH_CLIENT_SECRET"),
  redirect_uri = "https://app.example/callback",
  scopes = c("calendar.read", "calendar.write", "contacts.read"),
  token_targets = list(
    calendar = list(
      resource = "https://calendar.example/",
      scopes = c("calendar.read", "calendar.write"),
      required_scopes = "calendar.read",
      resource_ids = "calendar_api"
    ),
    contacts = list(
      resource = "https://contacts.example/",
      scopes = "contacts.read",
      resource_ids = "contacts_api"
    )
  ),
  default_token_target = "calendar",
  resource_bases = c(
    calendar_api = "https://calendar.example/v1",
    contacts_api = "https://contacts.example/v1"
  )
)
```

Several targets require an explicit default before sign-in. It determines code
redemption and later default accessors. With one target, `oauth_client()` selects
it automatically. Resource indicators and HTTP destinations are different
concepts: `resource_ids` explicitly authorizes a target's use at a named
`resource_bases` destination. Equal names never create an implicit association.
Omit HTTP bases and associations if the application only exports bearer tokens
to another library.

Configure at most 16 targets. The complete authorization has a shared limit of
128 distinct scope tokens and 8192 scope bytes, excluding separating spaces.
OIDC scopes count toward these limits. Configuration, acquired permissions
(including Microsoft `.default` expansion), restoration and reauthorization
enforce the same aggregate limits. Reduce the declared permissions or use
separate authorizations when the combined set would exceed them.
Managed transaction contexts support this full budget, including repeated limits
across targets, and are validated before reauthorization ends the current connection.
The independent callback and encrypted-state size limits still apply; a large
declaration can require larger state limits even when its scope count fits.

RFC 8707 authorization requests carry all declared resources and the client's
scopes. Code redemption and refresh carry one selected resource and its scopes.
The server determines which permissions apply to each resource; scopes in RFC
8707 are not independently paired with individual resources in the authorization
request. See [RFC 8707](https://www.rfc-editor.org/rfc/rfc8707.html).

Declare OIDC scopes at client level, outside the targets' API scope lists.
RFC 8707 mode recognizes `openid`, `profile`, `email`, `address`, `phone`, and
`offline_access`. All except `offline_access` can also be client-level
`required_scopes`. Request `offline_access` in client `scopes` for refresh
consent; putting it in client or target `required_scopes` is a configuration
error. A refresh token can be issued without this value appearing in the
access-token scopes, and its presence in scopes does not establish that a
usable refresh credential exists.
See [OpenID Connect scope values](https://openid.net/specs/openid-connect-core-1_0.html#ScopeClaims).
Microsoft mode keeps its provider-specific set: `openid`, `profile`, `email`,
and `offline_access`. It does not treat bare `phone` or `address` as Microsoft
OIDC scopes.

## Use the same connection interface in either module

Obtain target connections from the module's `connection()` factory. The legacy
`oauth_connection(client, token_reactive)` wrapper rejects target clients because
the token alone does not contain their retained permission limits.

```{r, eval=FALSE}
server <- function(input, output, session) {
  auth <- oauth_module_server("auth", client)
  # Alternatively: auth <- oauth_connections_server("auth", manager)
  # and select auth$connection(connection_id), or auth$connection() when unique.

  contacts <- shiny::reactive({
    connection <- shiny::req(auth$connection())
    token <- connection$access_token("contacts.read", target = "contacts")
    existing_contacts_sdk(access_token = token)$list_contacts()
  })

  output$can_edit_calendar <- shiny::renderText({
    connection <- shiny::req(auth$connection())
    connection$has_scopes("calendar.write", target = "calendar")
  })

  shiny::observeEvent(input$save_event, {
    connection <- shiny::req(auth$connection())
    connection$request(
      "calendar_api", "events", method = "POST",
      required_scopes = "calendar.write", target = "calendar", refresh = TRUE,
      configure = function(req) httr2::req_body_json(req, list(title = input$title))
    )
  })
}
```

The first contacts access can acquire its token even though `$has_scopes()` was
FALSE before that target had evidence. `$targets()` returns target names,
acquisition status and granted scope evidence, never credentials. Operation
`required_scopes` checks do not change the acquisition request. An explicitly
partial response remains that target's ceiling on later refreshes.

`connection$identity()` returns selected fields from the last validated OIDC
identity while the authorization remains available in the owning session. It
can still succeed after the primary access token expires or its current scopes
no longer include `openid`. This snapshot does not establish fresh user
authentication or current API access. Use `has_scopes()` and `access_token()` or
`request()` to check the selected target's permissions and token lifetime.

Each ceiling includes the OIDC scopes granted to that target. Explicit refresh
narrowing uses exactly the supplied scopes, so retain any OIDC permissions still
needed, for example `scopes = c("calendar.read", "openid")`. Required scopes
must remain present. Optional identity scopes such as `email` and `profile`
are not added back on refresh or restoration. Requested `offline_access` is
retained separately from access-token permission evidence: providers can omit
it from token scopes while issuing a refresh token. This preserves refresh
capability on reauthorization without reporting `offline_access` as a granted
access-token permission. Explicitly removing it from a refresh request removes
it from that target's retained limit as well. Reauthorization
requests the union of the retained target limits; an OIDC scope still retained
by another target can therefore remain in that authorization request.

OIDC reauthorization always requests `openid` for the new validated login,
including code redemption for the default target. This mandatory protocol scope
is restored in that new target's request limit even if the previous access-token
response omitted it or a refresh removed it. Removed API permissions and optional
OIDC scopes (`profile`, `email`, `address`, `phone`, `offline_access`) remain
excluded. Secondary targets keep their retained limits.

For RFC 8707, `connection$refresh(scopes = "calendar.read", target = "calendar")`
deliberately narrows calendar permissions. Acquiring contacts does not restore calendar
write access. Module `reauthorize()` retains all target ceilings and replaces
the connection ID; the manager takes an ID for this operation. To request a new
authorization with the original configuration after deliberate narrowing, use
the normal new-login flow rather than reauthorization of the restricted grant.

The cache and refresh credential belong to one authorization. Async accessors
for the same operation share work; different target acquisitions queue and read
the most recently committed refresh credential. A synchronous caller reports
`refresh_pending` while async work owns the credential. Safe target rejections
leave other entries usable. An ambiguous shared refresh outcome can invalidate
the whole authorization. Independent authorizations with equal refresh-token
bytes still follow the manager's conservative alias-invalidation rules.

With `indefinite_session = TRUE`, the single module retains cached target access
tokens after an ambiguous refresh failure, but retires the shared refresh
credential. A cached token can still be exported while it meets the operation's
lifetime and scope requirements. Forced refresh, expired tokens, and targets
not yet acquired require a new authorization; they cannot reuse the retired
credential. `token_stale` flags the failed refresh even while cached tokens are
usable. The manager has no indefinite-session exception.

The single module keeps a target authorization available for acquisition when
one access token expires. Its session and configured authorization-age limits
still apply. Its `token_stale` flag reports an expired primary token, and
refreshing a secondary target does not clear that flag. Its `authenticated`
flag tracks that retained authorization;
connection methods check the selected token's lifetime and permissions before
use. Recoverable target failures are returned to the caller without changing
the module's authentication or starting a new login. A successful token rotation
preserves the connection ID; logout,
replacement and session end invalidate old references and late completions.
Targets are stored inside the manager's existing encrypted credential envelope.

Single-module logout and configured session-end revocation remove local access
before contacting the provider. Remote target cleanup is best effort, with a
shared ten-second deadline and a two-second timeout per request without retries.
The shared refresh credential is attempted first. With `async = TRUE`, one
worker handles the batch; time spent waiting for a worker also counts toward
the deadline. Targets remaining after the deadline are not attempted.

`$request(refresh = TRUE)` is synchronous. It acquires before sending and
disables generic HTTP retries for that call. It never refreshes or replays an
application request after a resource error. DPoP nonce challenges retain the
existing bound-transport handling. Export through `$access_token()` remains
bearer-only; use `$request()` for DPoP/mTLS tokens. Writes belong in input-event
handlers and require another user action after an error.

## Microsoft configuration

```{r, eval=FALSE}
provider <- oauth_provider_microsoft(
  tenant = Sys.getenv("MICROSOFT_TENANT_ID"),
  userinfo_required = FALSE
)
client <- oauth_client(
  provider, Sys.getenv("OAUTH_CLIENT_ID"), Sys.getenv("OAUTH_CLIENT_SECRET"),
  redirect_uri = "https://app.example/callback",
  scopes = c(
    "openid", "offline_access",
    "https://graph.microsoft.com/Calendars.Read",
    "api://your-contacts-api/Contacts.Read"
  ),
  token_targets = list(
    calendar = list(
      resource = "https://graph.microsoft.com",
      scopes = "https://graph.microsoft.com/Calendars.Read"
    ),
    contacts = list(
      resource = "api://your-contacts-api",
      scopes = "api://your-contacts-api/Contacts.Read"
    )
  ),
  default_token_target = "calendar"
)
```

The preset enables Microsoft's scope-based acquisition mode. Authorization
requests include the declared consent scopes; token requests select one API's
qualified scopes plus its retained OIDC scopes (initially the configured OIDC
scopes), without an RFC 8707 `resource`
parameter. Disabling UserInfo avoids calling Graph with another API's token and
preserves ID-token validation. Resource identifiers are exact: where an
identifier ends in `/`, appending `/.default` produces a double slash.
See the [Microsoft code flow](https://learn.microsoft.com/en-us/entra/identity-platform/v2-oauth2-auth-code-flow).

For static consent, declare `resource/.default` as the target's sole API scope,
also in client `scopes`. Do not mix static and explicit API scope declarations.
Target `required_scopes` must name actual API permissions; `.default` is a consent
request, so it is rejected as a required access-token permission.
Microsoft static consent concerns the application's registered permissions for
all APIs; shinyOAuth sends one static scope at authorization and selects the
desired resource at token redemption. During reauthorization, the sealed target
limits still constrain token acquisition even if static consent remains needed
for an unacquired target.

`.default` is never treated as proof of individual permissions. Microsoft target
responses must include `scope`; returned short permission names are qualified
only within the selected Microsoft resource. Short names containing `:` or `/`
must match that target's declared scopes, required permissions, or retained
permissions from an earlier grant. URI-shaped names remain resource-qualified;
unknown punctuated short names must be returned in fully qualified form.
Token evidence records the actual
returned permissions. Microsoft API permission comparisons ignore ASCII case
(for example, `User.Read` and `user.read`); resource identifiers remain exact.
Ordinary OAuth scopes remain case-sensitive. Local operation limits retain only permissions within the
previous limit (expanding `.default` to its initial granted permissions), so a
later response cannot restore a locally removed operation. See [Microsoft scopes and static consent](https://learn.microsoft.com/en-us/entra/identity-platform/scopes-oidc)
and [refresh credentials](https://learn.microsoft.com/en-us/entra/identity-platform/refresh-tokens).

Entra can return all previously consented API permissions for the resource,
even when the request asks for fewer. Such responses are accepted without
enlarging the local operation limit. `$has_scopes()`, `$request()` and operation
requirements on `$access_token()` still enforce that limit, including during
refresh and after restoration. `$targets()` reports actual granted scopes, which
can therefore exceed the permissions available through these checks.

An exported bearer token still carries its full provider-issued permissions;
shinyOAuth cannot restrict what an external SDK does with it. Explicit Microsoft
API narrowing through `$refresh(scopes = ...)` fails before sending credentials,
with reason `unsupported_scope_narrowing`. The existing authorization stays
usable. Use a separate app registration or change provider consent when you need
a token with fewer API permissions. See [Microsoft consent lifetime](https://learn.microsoft.com/en-us/entra/msal/javascript/browser/resources-and-scopes#consent-lifetime).

These modes are covered by local HTTP contract tests, including rotation and
scope selection. Those tests do not establish compatibility with every RFC 8707
deployment or replace validation of your Microsoft registration and consent.
Discovery alone does not enable targets. SMART target switching, provider
plugins, distributed stores and incremental consent are outside this interface.

For SDK callbacks, event-bound result ownership and background tasks, see
`vignette("external-integrations")`.
